---
title: "TundraBUZZ Model Reports"
author: "Alexandre Beauchemin"
date: "2025-04-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Mismatch models


## 3.1.2. 

#### **Regression Coefficients**

| **Parameter**                     | **Estimate** | **Est.Error** | **95% CI**      | **Rhat** | **Bulk_ESS** | **Tail_ESS** |
|:----------------------------------|-------------:|--------------:|:---------------:|---------:|-------------:|-------------:|
| **Intercept**                      |     4.69     |      0.96     | [2.86, 6.62]    |     1.00 |     10274     |      8099     |
| **peak_flowering_numeric**         |     1.12     |      0.09     | [0.95, 1.29]    |     1.00 |     11216     |      7944     |
| summer_GDD0_100_c              |    -0.30     |      0.30     | [-0.88, 0.29]   |     1.00 |     11224     |      8264     |
| peak_flowering_numeric:summer_GDD0_100_c |  0.31  |      0.50     | [-0.65, 1.28]   |     1.00 |     11783     |      8333     |

#### **Further Distributional Parameters**

| **Parameter**                     | **Estimate** | **Est.Error** | **95% CI**      | **Rhat** | **Bulk_ESS** | **Tail_ESS** |
|:----------------------------------|-------------:|--------------:|:---------------:|---------:|-------------:|-------------:|
| **sigma**                          |     4.76     |      0.95     | [3.27, 6.99]    |     1.00 |      9706     |      8238     |



## 3.1.3. Binomial model of overlap probability

#### **Regression Coefficients**

| **Parameter**               | **Estimate** | **Est.Error** | **95% CI**       | **Rhat** | **Bulk_ESS** | **Tail_ESS** |
|:----------------------------|-------------:|--------------:|:----------------:|---------:|-------------:|-------------:|
| **Intercept**                |     1.20     |      0.32     | [0.59, 1.88]     |     1.00 |      8914     |      7848     |
| summer_GDD0_scaled       |    -0.66     |      0.38     | [-1.41, 0.08]    |     1.00 |      8473     |      7423     |



## 3.2. Drivers of bumblebee activity


#### **Correlation Structures**

| **Parameter**        | **Estimate** | **Est.Error** | **95% CI**      | **Rhat** | **Bulk_ESS** | **Tail_ESS** |
|:---------------------|-------------:|--------------:|:---------------:|---------:|-------------:|-------------:|
| ar[1]            |     0.44     |      0.35     | [-0.42, 0.91]   |     1.00 |       908     |       2337    |
| **sderr**            |     0.52     |      0.21     | [0.13, 0.98]    |     1.00 |       633     |       443     |

#### **Multilevel Hyperparameters** (Location-specific effects, number of levels: 9)

| **Parameter**                   | **Estimate** | **Est.Error** | **95% CI**      | **Rhat** | **Bulk_ESS** | **Tail_ESS** |
|:---------------------------------|-------------:|--------------:|:---------------:|---------:|-------------:|-------------:|
| **sd(Intercept)**                |     0.37     |      0.35     | [0.02, 1.18]    |     1.00 |      1674     |      1532     |
| **sd(mean_temp_z)**              |     0.25     |      0.20     | [0.01, 0.73]    |     1.00 |      2348     |      3564     |
| **sd(daily_nectar_sugar_mg_z)**  |     0.45     |      0.30     | [0.02, 1.15]    |     1.00 |      1587     |      2852     |
| cor(Intercept, mean_temp_z)  |    -0.04     |      0.49     | [-0.88, 0.85]   |     1.00 |      4978     |      5592     |
| cor(Intercept, daily_nectar_sugar_mg_z) |    -0.37 |      0.47     | [-0.96, 0.71]   |     1.00 |      1982     |      3261     |
| cor(mean_temp_z, daily_nectar_sugar_mg_z) |  0.13  |      0.48     | [-0.82, 0.90]   |     1.00 |      4601     |      5560     |

#### **Regression Coefficients**

| **Parameter**                     | **Estimate** | **Est.Error** | **95% CI**      | **Rhat** | **Bulk_ESS** | **Tail_ESS** |
|:----------------------------------|-------------:|--------------:|:---------------:|---------:|-------------:|-------------:|
| **Intercept**                     |     5.00     |      0.42     | [4.08, 5.60]    |     1.00 |      2293     |      1389     |
| **mean_temp_z**                   |     0.57     |      0.19     | [0.22, 0.98]    |     1.00 |      4669     |      3687     |
| daily_nectar_sugar_mg_z      |     0.27     |      0.24     | [-0.15, 0.81]   |     1.00 |      2965     |      3345     |
| **mean_wind_speed_z**             |    -1.15     |      0.13     | [-1.41, -0.89]  |     1.00 |      3230     |      4533     |
| mean_stn_press_k_pa_z         |     0.18     |      0.13     | [-0.08, 0.44]   |     1.00 |      6536     |      6351     |
| avg_air_rh_1m_z               |     0.28     |      0.17     | [-0.06, 0.61]   |     1.00 |      6987     |      5953     |
| **day_length_hours_z**            |    -1.14     |      0.37     | [-1.84, -0.40]  |     1.00 |      7677     |      6044     |
| cloud_cover_pct_z             |     0.03     |      0.15     | [-0.26, 0.31]   |     1.00 |      5835     |      5122     |

#### **Further Distributional Parameters**

| **Parameter**                     | **Estimate** | **Est.Error** | **95% CI**      | **Rhat** | **Bulk_ESS** | **Tail_ESS** |
|:----------------------------------|-------------:|--------------:|:---------------:|---------:|-------------:|-------------:|
| **shape**                         |     1.02     |      0.32     | [0.72, 1.60]    |     1.00 |       776     |       550     |

### TEMP ACTIVITY

```{r aic-comparison, echo=FALSE, message=FALSE, results='asis'}
library(knitr)

# AIC comparison table
df_comparison <- data.frame(
  df = 2:10,
  AIC = c(5771.070, 5771.100, 5753.791, 5757.449, 5755.492, 5753.873, 5754.502, 5755.753, 5757.876)
)

# Highlight the lowest AIC
df_comparison$Note <- ifelse(df_comparison$AIC == min(df_comparison$AIC), "Lowest AIC", "")

# Render table
kable(df_comparison,
      caption = "AIC Comparison Across Model Degrees of Freedom",
      digits = 3,
      align = c("r", "r", "l"))
```

## Model outputs

```{r lmer-spline-summary-full, echo=FALSE, message=FALSE}
library(knitr)

# REML criterion
reml_criterion <- 5698.5

# Scaled residuals
scaled_resid <- data.frame(
  Statistic = c("Min", "1Q", "Median", "3Q", "Max"),
  Value = c(-1.6580, -0.5513, -0.2105, 0.2277, 6.6206)
)

# Random effects
random_effects <- data.frame(
  Group = c("location_id (Intercept)", "Residual"),
  Variance = c(1761, 11695),
  Std_Dev = c(41.96, 108.14)
)

# Fixed effects
fixed_effects <- data.frame(
  Term = c("(Intercept)", "Spline 1", "Spline 2", "Spline 3", "Spline 4"),
  Estimate = c(-36.46, 59.90, 185.64, 302.97, 14.91),
  Std_Error = c(30.21, 26.83, 26.58, 67.67, 30.23),
  DF = c(145.85, 456.48, 456.58, 457.61, 456.86),
  `t value` = c(-1.207, 2.233, 6.984, 4.477, 0.493),
  `Pr(>|t|)` = c(0.229, 0.026, 1.02e-11, 9.56e-06, 0.622)
)

# Correlation of fixed effects
corr_fixed <- data.frame(
  Term1 = c("Spline 1", "Spline 2", "Spline 3", "Spline 4", "Spline 2", "Spline 3", "Spline 4", "Spline 3", "Spline 4", "Spline 4"),
  Term2 = c("(Intercept)", "(Intercept)", "(Intercept)", "(Intercept)", "Spline 1", "Spline 1", "Spline 1", "Spline 2", "Spline 2", "Spline 3"),
  Correlation = c(-0.731, -0.521, -0.857, -0.109, 0.236, 0.709, 0.325, 0.552, -0.321, 0.179)
)

# Output all tables
cat("### REML Criterion\n")
cat("REML criterion at convergence: **", reml_criterion, "**\n\n")

cat("### Scaled Residuals\n")
kable(scaled_resid, digits = 3)

cat("\n### Random Effects\n")
kable(random_effects, digits = 2)

cat("\n### Fixed Effects\n")
kable(fixed_effects, caption = "Fixed Effects Summary for Linear Mixed-Effects Model", digits = 3)

cat("\n### Correlation of Fixed Effects\n")
kable(corr_fixed, caption = "Pairwise Correlation of Fixed Effects", digits = 3)
```



### MICROCLIMATE CLUSTERING - PERMANOVA

```{r permanova-microclimate, echo=FALSE, message=FALSE, results='asis'}
library(knitr)

# Create the PERMANOVA results table
permanova_results <- data.frame(
  Term = c("Model", "Residual", "Total"),
  Df = c(2, 7, 9),
  SumOfSqs = c(0.60287, 0.86004, 1.46291),
  R2 = c(0.4121, 0.5879, 1.0000),
  F = c(2.4534, NA, NA),
  `Pr(>F)` = c(0.005, NA, NA),
  Signif = c("**", "", "")
)

# Render the table
kable(permanova_results,
      caption = "PERMANOVA Results Testing Microclimate Effects on Community Composition (999 Permutations)",
      digits = c(0, 0, 5, 4, 4, 3),
      align = c("l", "r", "r", "r", "r", "r"))
```

# Betadisper ANOVA

```{r dispersion-anova, echo=FALSE, message=FALSE, results='asis'}
library(knitr)

# Create the ANOVA table
dispersion_anova <- data.frame(
  Term = c("Groups", "Residuals"),
  Df = c(2, 7),
  SumSq = c(0.013992, 0.152834),
  MeanSq = c(0.0069961, 0.0218334),
  F = c(0.3204, NA),
  `Pr(>F)` = c(0.7359, NA)
)

# Render the table
kable(dispersion_anova,
      caption = "Test for Homogeneity of Multivariate Dispersions (ANOVA on betadisper distances)",
      digits = c(0, 0, 6, 7, 4, 4),
      align = c("l", "r", "r", "r", "r", "r"))
```

# Pairwise results

```{r pairwise-permanova, echo=FALSE, message=FALSE, results='asis'}
library(knitr)

# Pairwise PERMANOVA results
pairwise_results <- data.frame(
  Group1 = c("Cool", "Cool", "Moderate"),
  Group2 = c("Moderate", "Warm", "Warm"),
  F_value = c(2.35, 3.27, 1.75),
  R2 = c(0.320, 0.396, 0.305),
  p_value = c(0.026, 0.019, 0.1)
)

# Add significance stars
pairwise_results$Signif <- cut(pairwise_results$p_value,
  breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
  labels = c("***", "**", "*", ".", "")
)

# Print the table
kable(pairwise_results,
      caption = "Pairwise PERMANOVA Results Between Microclimate Groups",
      digits = c(0, 0, 2, 3, 3),
      align = c("l", "l", "r", "r", "r", "c"))
```




### SCALING ANALYSIS - PERMANOVA TABLES

```{r PERMANOVA}
# Create the table data
adonis_results <- data.frame(
  Term = c("Model", "Residual", "Total"),
  Df = c(1, 18, 19),
  SumOfSqs = c(0.1921, 3.6292, 3.8214),
  R2 = c(0.05028, 0.94972, 1.00000),
  F = c(0.9529, NA, NA),
  `Pr(>F)` = c(0.46, NA, NA)
)

# Print the table with formatting
kable(adonis_results, caption = "PERMANOVA Results (adonis2) with 999 Permutations",
      digits = c(0, 0, 4, 5, 4, 2),
      align = c("l", "r", "r", "r", "r", "r"))
```

### Homogeneity test

```{r dispersion-test, echo=FALSE, message=FALSE, results='asis'}
library(knitr)

# Create the table data
dispersion_results <- data.frame(
  Term = c("Groups", "Residuals"),
  Df = c(1, 18),
  SumSq = c(0.02057, 0.36470),
  MeanSq = c(0.020575, 0.020261),
  F = c(1.0155, NA),
  NPerm = c(999, NA),
  `Pr(>F)` = c(0.327, NA)
)

# Print the table with formatting
kable(dispersion_results,
      caption = "Test for Homogeneity of Multivariate Dispersions (999 Permutations)",
      digits = c(0, 0, 5, 6, 4, 0, 3),
      align = c("l", "r", "r", "r", "r", "r", "r"))
```











